---
title: "Brazil Rent Report - Data Analysis for Business 2023"
author: "Eduardo Mosca, Chakir El Arrag, Iraj Nayebkhil"
date: "`r Sys.Date()`"
output: pdf_document
editor_options: 
  markdown: 
    wrap: sentence
geometry: margin=1in
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

In this report, we look at the dataset for rent prices in Brazil, specifically along the south border, looking for specifics of variables and characteristics which might help paint the picture for our (hypothetical) employer's entry in the market.
We now take a look at the dataset and its variables.

```{r,echo = FALSE,warning=FALSE,message=FALSE}
# Importing the libraries needed 
library(gam)
library(xgboost)
library(tidyverse)
library(glmnet)
library(caret)
library(ggplot2)
library(ggrepel)
library(ggpubr)
library(dplyr)
library(plyr)
library(maps)
library(cowplot)
library(MASS)
library(rpart)
library(rpart.plot)
library(gglasso)
library(factoextra)
library(cluster)


# Importing the dataset
Data = read.csv("C:/Users/user2/Desktop/dab2/BrazHousesRent.csv", sep = ",", dec = ".", header = T, colClasses = "character")

```

# 1 - Overview of the dataset

Of the 12 variables on hand, five of them continuous (fire insurance, property tax, HOA fee, rent amount and area), three are categorical(animal accepted,apartment furnished, and city where property is located).
The remaining four are numerical discrete, they represent the floor number and the number of rooms, bathrooms, and parking spots.
We have a rough dataset of 10692 observations, not a lot but enough to extract valuable info.
We start cleaning the dataset: there are no specific NA's but the floor has "-" values.
By calling anyDuplicated we also notice some duplicate rows, we want to remove them with "unique()".

```{r, echo = FALSE,warning=FALSE,message=FALSE}
# Remiving duplicate rows
Data <- unique(Data)
```

The floor column has a '-' value which is pretty occurent in our dataset.

```{r}
# Counting the occurences
count_zero <- sum(Data$floor == "-")
print(paste("The number of observations with floor '-' is:", count_zero))
```

We have 2369 observations with floor "-" on our dataset, interestingly the minimum value in the floor column is 1, which suggests that "-" would represent apartments in the ground floor or normal houses that are not in a condominium.
To be sure, we extract from the dataset the HOA (Homeowners Association Tax) for houses on the '-' floor.

```{r}
# Counting observations with floor '-' and hoa 0, we also show their proportion
floor_abs <- subset(Data, floor == "-", select = hoa..R..)
floor_abs[,1] = as.numeric(floor_abs[,1])
count_zero <- sum(floor_abs[, 1] == 0)
print(paste("Number of observations with floor '-' and HOA = 0:", count_zero))
print(paste("Proportion of observations with floor '-' and HOA = 0:", count_zero/length(floor_abs[, 1])))
```

As we can see, 85,06% of the observations has an HOA of 0, which strengthens our hypothesis since owners of normal houses don't have an HOA tax to pay because the houses are not situated in a condominium.
The rest of the occurences with HOA higher than 0, can be justified by the presence of some condominiums with houses on the ground floor, not very frequent but this would justify why they're paying the HOA tax while at the same time being at floor '-'.
Since the data is compatible with our hypothesis, we'll proceed by replacing '-' by 0.

```{r}
# Replace '-' by 0 in the floor column
Data[Data == "-"] <- 0
```

Next, we set the variables to be either numeric or factors.

```{r}
categ <- c("city","animal","furniture")
cols <- colnames(Data)
for(i in 1:ncol(Data)){
  if (cols[i] %in% categ){
    Data[,i] = as.factor(Data[,i])
  }else{
    Data[,i] = as.numeric(Data[,i])
  }
}
```

# 2 - Looking at the response variable

We can look at the geographical position of the cities we have info on, and visualize which ones are more expensive on average.

```{r}
# Function that computes the average rent amount for a given city
average_rent_price <- function(City) {
  city_rent_amounts <- subset(Data, city == City, select = rent.amount..R..)
  average_rent <- mean(city_rent_amounts$rent.amount..R..)
  return(average_rent)
}
```

```{r fig2, fig.height = 4, fig.width = 6, fig.align = "center",,warning=FALSE,message=FALSE,echo=FALSE}
city_data <- data.frame(
  City = c("Belo Horizonte", "Campinas", "Porto Alegre", "Rio de Janeiro", "Sao Paulo"),
  Latitude = c(-19.9167, -22.9071, -30.0346, -22.9068, -23.5505),
  Longitude = c(-43.9345, -47.0632, -51.2177, -43.1729, -46.6333),
  Average_Rent = sapply(c("Belo Horizonte", "Campinas", "Porto Alegre", "Rio de Janeiro", "São Paulo"), 
                        function(x) average_rent_price(x))
)

city_data$Alpha <- (city_data$Average_Rent -
                      min(city_data$Average_Rent)) / (max(city_data$Average_Rent) 
                                                      - min(city_data$Average_Rent))

# Plot the geospatial data for Brazil only
map_brazil <- map_data("world", region = "Brazil")

# Create the plot
ggplot() +
  geom_polygon(data = map_brazil, aes(x = long, y = lat, group = group), fill = "lightgray", color = "white") +
  geom_label_repel(data = city_data, aes(x = Longitude, y = Latitude, label = City), color = "black", size = 3,
                   box.padding = 0.5, point.padding = 0.2, force = 1, segment.color = "transparent") +
  geom_point(data = city_data, aes(x = Longitude, y = Latitude, color = Average_Rent), alpha = 0.8, size = 5) +
  labs(title = "Average Rent in Different Cities in Brazil",
       x = "Longitude",
       y = "Latitude") +
  scale_color_gradient(low = "blue", high = "red") +
  theme_minimal()

```

As we can see Sao Paulo tends to be the most expensive place to rent a house.

A high correlation is observed between fire insurance and rent, suggesting that fire insurance could be a valuable variable for modeling.

```{r}
cor(Data$fire.insurance..R..,Data$rent.amount..R..)
```

### Defining our objective

Our objective is to *Build a predictive model and find out the rent amount according to the house specifics*, by using regression methods on the response variable rent.amount..R..

# 3 - Lower Dimensional Models

When thinking about low dimension models, we must understand which ones might apply to our use case, yielding information that is actually valuable.
We need to better understand relations between the response and some variables.
Its valuable to know how much the rent increases as the number of rooms or area increases.
We first implement a Log-Linear regression model to see what percentage increase there is when the area goes up by one unit:

```{r}
fit1 <- lm(log(rent.amount..R..) ~ area,data = Data)
summary(fit1)$coefficients
```

By looking at the coefficients, we see the expected % increase following an increase in the area by 1 unit.
We also want to model the reasoning of a person interested in renting an apartment in Brazil via decision tree, by considering the main concerns when house shopping.
This will give us insight on the price levels according to specifics:

```{r, fig0, fig.height = 3, fig.width = 4, fig.align = "center",,warning=FALSE,message=FALSE}
dtree <- rpart(rent.amount..R.. ~ rooms + furniture + bathroom, data = Data, method = "anova")
rpart.plot(dtree)
```

The specifics leading to higher rent are to the right, while following the left-most branch we get cheaper rent.

# 4 - Getting the data ready

We start by removing outliers in the continuous variables using the IQR.
For discrete numerical variables, we give a look at the boxplots and we remove outliers manually.

```{r}
# Continuous variables
cont <- c("area","fire.insurance..R..","property.tax..R..","hoa..R..", "floor")

#Removing outliers of continuous variables using IQR
for(i in 1:ncol(Data)){
  if (cols[i] %in% cont){
    Q3 <- quantile(Data[,i], .75)
    IQR <- IQR(Data[,i])
    Data <- subset(Data, Data[,i]< (Q3 + 3.5*IQR))
  }
}

# Boxplots for discrete variables
# discrete <- c("rooms", "floor", "parking.spaces", "bathroom")
# for(i in 1:length(discrete)){
#   boxplot(Data[, discrete[i]], xlab = discrete[i])
# }

#Removing outliers from discrete variables
Data <- subset(Data, floor < 40)
Data <- subset(Data, rooms < 10)
Data <- subset(Data, bathroom < 7)
```

Then, we turn the variables animal and furniture to 0s and 1s, the variables in the original dataset take a binary value encoded as a character (ex. for animal we have acept and not acept)

```{r, warning=FALSE, message=FALSE}
# Turning variables animal and furniture to 0s and 1s

#Animal
vec <- Data$animal
vec <- as.character(vec)
vec[vec == "acept"] <- 1      # Replace "acept" by 1
vec[vec == "not acept"] <- 0       # Replace "not acept" by 0
vec <- as.factor(vec)
Data$animal <- vec

#Furniture
vec <- Data$furniture
vec <- as.character(vec)
vec[vec == "furnished"] <- 1      # Replace "furnished" by 1
vec[vec == "not furnished"] <- 0       # Replace "not furnished" by 0
vec <- as.factor(vec)
Data$furniture <- vec
```

Next, we split the dataset into training, validation and test sets, and we scale them using the mean and standard deviation of the training set.
We save main training set (training + validation) to work with it later for the test set prediction.

```{r,warning=FALSE,message=FALSE}
#Train test split
# We first split into training and test set, then we split the training set into a smaller training set and a validation set
set.seed(1)
trainrows <- createDataPartition(Data$rent.amount..R.., p=0.8, list=FALSE)
training_set <- Data[trainrows,] # This is saved and will be scaled later to work on the test set
d_test <- Data[-trainrows,]
trainrows <- createDataPartition(training_set$rent.amount..R.., p=0.8, list=FALSE)
d_train <- training_set[trainrows,]
d_val <- training_set[-trainrows,]

unscaled_rent_amounts <- d_test$rent.amount..R.. # Will be saved for later use
mean_tr <- mean(training_set$rent.amount..R..)
std_tr <- sd(training_set$rent.amount..R..)

# Scaling the data
# We make a function that scales the data in dataset using the mean and sd of dataset2
scale_data <- function(dataset, dataset2) {
  for(i in 1:ncol(dataset)){
    if (is.numeric(dataset[1,i])){
      dataset[,i] = scale(dataset[,i], center = mean(dataset2[,i]), scale = sd(dataset2[,i]))
    }
  }
  return(dataset)
} 

# Scaling the training, validation and test sets
d_val <- scale_data(d_val, d_train)
d_test <- scale_data(d_test, training_set)
d_train <- scale_data(d_train, d_train) #Training set last since we use its mean and sd to scale the val set
```

Finally, we encode categorical variables (We keep an unencoded version of the training and validation sets for later).

```{r,warning=FALSE,message=FALSE, echo=FALSE}
# Saving unencoded versions of training and validation sets
d_train_unenc <- d_train
d_val_unenc <- d_val

# We make a function to encode the categorical variables in a dataset
encode <- function(dataset, excluded=c()) {
  excluded_cols <- dataset[, names(dataset) %in% excluded]
  dataset <- dataset[,!names(dataset) %in% excluded]
  dmy <- dummyVars(" ~ .", data = dataset)
  dataset <- data.frame(predict(dmy, newdata = dataset))
  dataset <- cbind(dataset, excluded_cols)
  return(dataset)
}  

# Encoding the categorical variables in the training, validation and test sets
d_train <- encode(d_train, c("animal", "furniture"))
d_val <- encode(d_val, c("animal", "furniture"))
d_test <- encode(d_test, c("animal", "furniture"))

# Removing one city (presence of a house in this city would be represented by 0s in the other four city columns, so we can remove one city)
d_train <- subset(d_train, select = -c(city.Campinas))
d_val <- subset(d_val, select = -c(city.Campinas))
d_test <- subset(d_test, select = -c(city.Campinas))
```

# 5 - Testing the models' performances, Model Selection

## 5.1 - AIC and BIC

First we implement AIC and BIC stepwise selection for multiple regression models, then we evaluate the performance on the validation set using the following metrics: **MSE**, **RMSE** and **R Squared**.

```{r,echo=FALSE}
# Implemention of AIC and BIC stepwise selection for multiple regression models
full.model <- lm(rent.amount..R.. ~ ., data = d_train)
step.model.aic <- stepAIC(full.model, direction = "both", trace = 0)

step.model.bic <- stepAIC(full.model, direction = "both", trace = 0, k = log1p(nrow(d_train)))

predictions.aic <- predict(step.model.aic, newdata = d_val)
predictions.bic <- predict(step.model.bic, newdata = d_val)


# Models' performance on the validation set
aic_mse <- mean((predictions.aic - d_val$rent.amount..R..)^2)
bic_mse <- mean((predictions.bic - d_val$rent.amount..R..)^2)
aic_rmse <- sqrt(aic_mse)
bic_rmse <- sqrt(bic_mse)
aic_rsquared <- summary(step.model.aic)$r.squared
bic_rsquared <- summary(step.model.bic)$r.squared

#Data frame to store the evaluation metrics
table <- data.frame(
  Model = character(),  
  MSE = numeric(),
  RMSE = numeric(),
  R_squared = numeric()
)
table <- rbind(table, c("AIC", round(aic_mse, 5), round(aic_rmse, 5), round(aic_rsquared, 5)))
table <- rbind(table, c("BIC", round(bic_mse, 5), round(bic_rmse, 5), round(bic_rsquared, 5)))
colnames(table) <- c("Model", "MSE", "RMSE", "R_squared")

print(table)
```

Our two models tend to yield similar results, BIC tends to be slightly better, so we're gonna consider it over AIC.
We can also plot the residuals against the fitted values of each:

```{r fig7, fig.height = 3, fig.width = 7, fig.align = "center",echo=FALSE,warning=FALSE,message=FALSE}
bicplot <- ggplot(data.frame(Fitted = step.model.bic$fitted.values, Residuals = step.model.bic$residuals), aes(x = Fitted, y = Residuals)) +
  geom_point(color = "blue") + 
  theme_minimal() +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(title = "Res vs. Fitted BIC", x = "Fitted Values", y = "Residuals")
aicplot <- ggplot(data.frame(Fitted = step.model.aic$fitted.values, Residuals = step.model.aic$residuals), aes(x = Fitted, y = Residuals)) +
  geom_point(color = "blue") + 
  theme_minimal() +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(title = "Res vs. Fitted AIC", x = "Fitted Values", y = "Residuals")
ggarrange(aicplot,bicplot,nrow = 1,ncol = 2)
```

The two are pretty similar and for the most part the points hover around the 0 line, which indicats a good fit.
When getting a similar result for AIC and BIC we can assume to be striking a good balance between model complexity and goodness of fit.

## 5.2 - Lasso and Group-Lasso

Let's implement a lasso regression using glmnet and group-Lasso using gglasso.
We start directly from 10-fold cross validation, and then we evaluate the model on the validation set using the lambda.min and lambda.1se

```{r fig4, fig.height = 6, fig.width = 6, fig.align = "center",out.width=2,echo=FALSE,warning=FALSE,message=FALSE}
# Performing 10-fold cross validation
X = model.matrix(rent.amount..R.. ~ ., data=d_train)[,-1]
y = d_train$rent.amount..R..
cvlasso = cv.glmnet(x = X, y = y,nfolds = 10)
groups <- c(1,1,1,1,2,3,4,5,6,7,8,9,10,11)
glasso = cv.gglasso(x=X, y=y, group = groups,nfolds = 10)
```

```{r,echo=FALSE}
# Models' performance on the validation set using the lambda.min and lambda.1se

pen_val <- model.matrix(rent.amount..R.. ~ ., data=d_val)[,-1]
lassopredmin <- predict(cvlasso,pen_val, s = "lambda.min")
lmin_mse <- mean((lassopredmin - d_val$rent.amount..R..)^2)
lmin_rmse <- sqrt(lmin_mse)
lmin_Rsq <- cor(lassopredmin,d_val$rent.amount..R..)^2
lassopred1se <- predict(cvlasso,pen_val, s = "lambda.1se")
lse_mse <- mean((lassopred1se - d_val$rent.amount..R..)^2)

glpred1 <- predict(glasso,pen_val, s="lambda.1se")
gl1_MSE <- mean((glpred1 - d_val$rent.amount..R..)^2)
gl1_Rsq <- cor(glpred1,d_val$rent.amount..R..)^2
glpred2 <- predict(glasso,pen_val, s="lambda.min")
gl2_MSE <- mean((glpred2 - d_val$rent.amount..R..)^2)
gl2_Rsq <- cor(glpred2,d_val$rent.amount..R..)^2

```

The lambda.min values get a lower error on the validation set, so let's compare the two models using minimum lambda:

```{r,fig5, fig.height = 4, fig.width = 4,fig.align = "center",echo=FALSE,warning=FALSE,message=FALSE,}
#Data frame to store the evaluation metrics
table <- data.frame(
  Model = character(),  
  MSE = numeric(),
  RMSE = numeric(),
  R_squared = numeric()
)
table <- rbind(table, c("Lasso", round(lmin_mse, 5), round(sqrt(lmin_mse), 5), round(lmin_Rsq, 5)))
table <- rbind(table, c("grLasso", round(gl2_MSE, 5), round(sqrt(gl2_MSE), 5), round(gl2_Rsq, 5)))
colnames(table) <- c("Model", "MSE", "RMSE", "R_squared")
print(table)
```

The two models yield similar results, with Lasso having slightly lower errors and a higher R Squared value.

## 5.3 - GAM

In order to implement GAM, we create the smoothing terms for numerical and categorical variables (We're going to use the unencoded versions of the training and validation sets we saved earlier), then we create the GAM formula summing the smoothing terms and we feed it to the model for the fitting.

```{r}
# Smoothing terms for numerical variables
num_names = names(d_train_unenc)[d_train_unenc %>% map_lgl(is.numeric)]
num_names = num_names %>% 
  discard(~.x %in% c("rent.amount..R.."))
num_feat = num_names %>% 
  map_chr(~paste0("s(", .x, ", 10)")) %>%
  paste(collapse = "+")

# Smoothing terms for categorical variables
cat_feat = names(d_train_unenc)[d_train_unenc %>% map_lgl(is.factor)] %>% 
  paste(collapse = "+")

# GAM formula
gam_form = as.formula(paste0("rent.amount..R.. ~", num_feat, "+", cat_feat))

# Model fitting
fit_gam = gam(formula = gam_form, family = "gaussian", data = d_train_unenc)
```

```{r fig111,fig.height=3,fig.width=6, fig.align='center'}
ggplot(data.frame(Fitted = fit_gam$fitted.values, Residuals = fit_gam$residuals), aes(x = Fitted, y = Residuals)) +
  geom_point(color = "red") + 
  theme_minimal() +
  geom_hline(yintercept = 0, linetype = "dashed", color = "green") +
  labs(title = "Residuals vs Fitted Values", x = "Fitted Values", y = "Residuals")
```

In this case as well, most of the points hover around the 0 line, with most residuals being in the interval $\left [ -0.5, 0.5 \right ]$, which indicates a good fit.
Next, we predict on the validation set, and we evaluate the predictions using the metrics mentioned previously.
We save the results for the models comparison.

```{r}
# Predicting the validation set
predicted_values = predict(fit_gam, d_val_unenc) 

# Evaluation metrics
observed_values = d_val_unenc$rent.amount..R..
gam_mse = mean((predicted_values - observed_values)^2)
gam_rmse = sqrt(gam_mse)
gam_rsquared = cor(predicted_values, observed_values)^2
```

## 5.4 - XGBoost

Now we create an XGBoost model.
We start by splitting dependent and independent variables and transforming the training and validation sets to xgb.Dmatrices which will be used for training and prediction.

```{r, warning=FALSE,message=FALSE}
X_train <- d_train[, !(colnames(d_train) %in% c("rent.amount..R.."))]
y_train <- as.numeric(d_train$rent.amount..R..)

X_val <- d_val[, !(colnames(d_val) %in% c("rent.amount..R.."))]
y_val <- as.numeric(d_val$rent.amount..R..)

#Make sure the columns are numeric before making the xgb.DMatrix that XGBoost is gonna use
for(i in 1:ncol(X_train)){
  X_train[,i] = as.numeric(X_train[,i])
}

for(i in 1:ncol(X_val)){
  X_val[,i] = as.numeric(X_val[,i])
}

xgb_train <- xgb.DMatrix(data = as.matrix(X_train), label = y_train)
xgb_val <- xgb.DMatrix(data = as.matrix(X_val))
```

We use 10-fold cross validation for parameter tuning.
The parameters we want to tune are the following:
The number of boosting rounds `nrounds`, the maximum tree depth `max_depth` and the learning rate `eta`.

```{r, warning=FALSE,message=FALSE}
# Parameter grid
param_grid <- expand.grid(
  nrounds = c(100, 200, 300),
  max_depth = c(3, 4, 5),             
  eta = c(0.01, 0.05, 0.1),
  gamma = 0,
  colsample_bytree = 1,
  min_child_weight = 1,  
  subsample = 1   
)

# The other parameters that have a fixed value are included in the grid just because excluding them throws an error when we run cv, however we gave them a fixed value (default value of the model) so that we don't include them in the cross validation process
```

We put the parameters to tune and their respective values to test in a parameter grid.
Then, we use the parameter grid as input to perform 10-fold cross validation in order to get the best combination of values for the parameters.

```{r, warning=FALSE,message=FALSE}
# 10-fold cross-validation
xgb_model <- train(
  X_train, y_train, 
  method = "xgbTree",
  trControl = trainControl(method = "cv", number = 10),
  tuneGrid = param_grid,
  metric = 'RMSE',
  verbosity = 0
)
```

Finally, we fit the final XGBoost model with the best parameters we got, we proceed to predict on the validation set and we evaluate the prediction, We save the errors for the models' comparison.

```{r, warning=FALSE,message=FALSE}
# Model fitting
final_model <- xgboost(
  data = xgb_train, 
  nrounds = xgb_model$bestTune$nrounds,
  max_depth = xgb_model$bestTune$max_depth,
  eta = xgb_model$bestTune$eta,
  verbose = 0
)

# Evaluation of the model
predictions <- predict(final_model, newdata = xgb_val)
xgb_mse <- mean((predictions - y_val)^2)
xgb_rmse <- sqrt(xgb_mse)
xgb_rsquared <- 1 - (sum((y_val - predictions)^2) / sum((y_val - mean(y_val))^2))
```

## 5.5 - Model Selection

We make a table with the errors for the models tested until now

```{r}
#Data frame to store the evaluation metrics
table <- data.frame(
  Model = character(),  
  MSE = numeric(),
  RMSE = numeric(),
  R_squared = numeric()
)

table <- rbind(table, c("BIC", round(bic_mse, 5), round(bic_rmse, 5), round(bic_rsquared, 5)))
table <- rbind(table, c("LASSO", round(lmin_mse, 5), round(lmin_rmse, 5), round(lmin_Rsq, 5)))
table <- rbind(table, c("GAM", round(gam_mse, 5), round(gam_rmse, 5), round(gam_rsquared, 5)))
table <- rbind(table, c("XGBoost", round(xgb_mse, 5), round(xgb_rmse, 5), round(xgb_rsquared, 5)))
colnames(table) <- c("Model", "MSE", "RMSE", "R_squared")

print(table)
```

We notice that the model that performs the best is XGBoost, we're going to use it to predict on the test set.

# 6 - Prediction on the test set

We start by scaling the training set (training + validation) and we encode its categoricals, then we transform the training and test sets to xgb.Dmatrices.
Next, we fit the model and we make predictions on the test set without forgetting to evaluate our predictions (also for unscaled data for interpretability) and we estimate a 95% confidence interval for the error which will provide an interval estimate of the potential deviation of the predicted rent prices from the true rent prices.

```{r, warning=FALSE,message=FALSE}
# Scaling and encoding training set
training_set <- scale_data(training_set, training_set)
training_set <- encode(training_set, c("animal", "furniture"))
training_set <- subset(training_set, select = -c(city.Campinas))

# Splitting dependent and independent variables
X_train <- training_set[, !(colnames(training_set) %in% c("rent.amount..R.."))]
y_train <- as.numeric(training_set$rent.amount..R..)

X_test <- d_test[, !(colnames(d_test) %in% c("rent.amount..R.."))]
y_test <- as.numeric(d_test$rent.amount..R..)

# Making sure the columns are numeric
for(i in 1:ncol(X_train)){
  X_train[,i] = as.numeric(X_train[,i])
}
for(i in 1:ncol(X_val)){
  X_test[,i] = as.numeric(X_test[,i])
}

# Transforming the training and test sets to xgb.Dmatrices
xgb_train <- xgb.DMatrix(data = as.matrix(X_train), label = y_train)
xgb_test <- xgb.DMatrix(data = as.matrix(X_test))

# Model fitting
final_model <- xgboost(
  data = xgb_train, 
  nrounds = xgb_model$bestTune$nrounds,
  max_depth = xgb_model$bestTune$max_depth,
  eta = xgb_model$bestTune$eta,
  verbose = 0
)

# Predicting on the test set
predictions <- predict(final_model, newdata = xgb_test)
unscaled_predictions <- (predictions * std_tr) + mean_tr

# Evaluation of the performance
test_mse <- mean((predictions - y_test)^2)
test_rmse <- sqrt(xgb_mse)
test_rsquared <- 1 - (sum((y_test - predictions)^2) / sum((y_test - mean(y_test))^2))
unscaled_mse <- mean((unscaled_predictions - unscaled_rent_amounts)^2)
unscaled_rmse <- sqrt(unscaled_mse)
print(paste("MSE: ", round(test_mse, 5)))
print(paste("RMSE: ", round(test_rmse, 5), "| RMSE for unscaled predictions", round(unscaled_rmse, 5)))
print(paste("R Squared: ", round(xgb_rsquared, 5)))

# We get the 95% confidence interval for the error of our model
prediction_errors <- unscaled_predictions - unscaled_rent_amounts

# Mean and standard deviation of the prediction errors
mean_error <- mean(prediction_errors)
std_error <- sd(prediction_errors)

# 95% confidence interval for the error
confidence <- 0.95
z_value <- qnorm(1 - (1 - confidence) / 2)
lower_bound <- mean_error - z_value * std_error  # Lower bound of the confidence interval
upper_bound <- mean_error + z_value * std_error  # Upper bound of the confidence interval

cat("Confidence Interval (", confidence * 100, "%): [", lower_bound, ", ", upper_bound, "]\n")
```

Our findings indicate that the model is consistent and performs well on unseen data, as the low errors obtained in the test set closely align with those from the validation set.
The confidence interval we got represents the range within which the true error of predicting rent prices in real currency is expected to lie with 95% confidence, In other terms, the values within the interval represent the potential deviation of the predicted rent prices from the true (actual) rent prices. The predicted rent prices could be up to **654.36 Real** lower than the true rent prices on average and up to **630.64 Real** higher than the true rent prices on average, which is a very good result considering the range of rent amounts we're dealing with. Therefore, the model is capable to accurately predict rent prices for houses in Brazil's major cities.
This predictive capability can empower the company to assess potential returns on rental investments before committing to them.
By leveraging cost data and the model's reliable profit predictions, we can identify optimal investment opportunities.
This involves identifying houses with lower costs that can be rented out at higher prices, maximizing potential returns.
Additionally, the high R-squared value indicates that a large proportion of the variance is explained by the variables in our model.
In other terms, our model accurately captures and predicts how rental prices behave based on the selected independent variables. It means that there is a strong and consistent relationship between these variables and the rent price.
However, it is important to assess whether all independent variables participate equally in this relationship by evaluating each variables' importance.

```{r fig22, fig.width=6,fig.height=3,fig.align='center'}
#Feature importance scores
importance_scores <- xgb.importance(
  feature_names = colnames(X_train),
  model = final_model
)

#Plot feature importance
xgb.plot.importance(importance_matrix = importance_scores)
```

Based on the plot, we can see that the variable with the highest importance score is `fire.insurance..R..`. Most of the other variables have significantly lower importance scores compared to it.
One of the notable variables with high importance is `city`, which aligns with our previous observations from the geospatial plot in which we found that average rent prices varied significantly across different cities.
Porto Alegre and São Paulo emerged as the cities with the highest importance scores.
Indeed, the geospatial plot demonstrated that Porto Alegre had the lowest average rent prices, while São Paulo had the highest.
Consequently, knowing the city of a house provides valuable information for predicting rent prices.
Two other significant variables are `floor` and `hoa..R..`.
Surprisingly, the area of the house and the number of rooms do not hold as much importance as one might typically expect.
This could be because the location of the house plays a more important role.
For instance, it might be more cost-effective to rent a larger house in Porto Alegre compared to a smaller apartment in the center of São Paulo.

The variable `fire.insurance..R..` has a very high importance, let's make a scatterplot to see if there's any linear/non-linear relationship with the response.

```{r fig70, fig.width=6, fig.height=4, fig.align='center'}
plot(Data$rent.amount..R.., Data$fire.insurance..R.., 
     xlab = "Rent Amount", ylab = "Fire Insurance",
     main = "Scatter Plot of Rent Amount vs Fire Insurance")
```

There is a strong linear relationship between fire insurance and rent amount.
This is not strange since fire insurance premiums are often based on the value of the insured property and the associated risk factors.
In this case, higher rent prices are typically associated with properties of higher value.
Properties with higher values are likely to have higher replacement costs and, consequently, higher fire insurance premiums.
It is worth to mention that insurance premiums can be also influenced by the location of the property.
Indeed, from the plot we see multiple lines, we can assume that each line could be representing a city. Let's investigate further.

```{r fig77, fig.width=6, fig.height=3, fig.align='center'}
# Create a scatter plot with different colors for each city
ggplot(Data, aes(x = rent.amount..R.., y = fire.insurance..R.., color = city)) +
  geom_point() +
  labs(title = "Rent Amount vs. Insurance by City") +
  xlab("Rent Amount") +
  ylab("Insurance")
```

After analyzing the plot, we observed that the lines representing cities with lower average rent amounts had steeper slopes.
This implies that for properties of the same value in both cities, the property in the city with the lower average rent amount would incur higher fire insurance costs.
The reason behind this is that even though the properties have the same value, the ones located in the city with the lower average rent amount tend to be larger in terms of area and number of rooms.
Consequently, they have higher replacement costs, which justifies the higher insurance premiums.
Furthermore, we noticed that each city displayed two distinct lines, we believe that there is another influential factor impacting both fire insurance costs and rent amounts.
We think it combines the two variables `floor` and `hoa..R..` since they were identified as important for our model.
This crucial factor is whether the house is located within a condominium or not.
When a house has a floor value of 0, it indicates that it is either situated on the ground floor of a condominium or it is an independent house not associated with any condominium.
However, what distinguishes between these two scenarios is the presence of the HOA.
If the HOA value is 0, it means that the house in question is an independent house.
On the other hand, if the HOA value is greater than 0, it indicates that the house is part of a condominium.
To investigate further, we can create a plot for Belo Horizonte, differentiating between houses in condominiums and independent houses.

```{r fig78 , fig.width=6, fig.height=3, fig.align='center'}
# Filtering independent houses and apartments in Belo Horizonte
hoa_zero_data <- subset(Data, Data$city == "Belo Horizonte" & Data$hoa..R.. == 0 & Data$floor == 0)
hoa_nonzero_data <- subset(Data, Data$city == "Belo Horizonte" & Data$hoa..R.. != 0)
hoa_data <- rbind(transform(hoa_zero_data, Group = "Independent"),
                  transform(hoa_nonzero_data, Group = "Condominium"))

# Scatterplot differenciating both groups
ggplot(hoa_data, aes(x = rent.amount..R.., y = fire.insurance..R.., color = Group)) +
  geom_point() +
  labs(title = "Rent Amount vs. Insurance in Belo Horizonte",
       subtitle = "") +
  xlab("Rent Amount") +
  ylab("Insurance")
```

The plot indeed confirms our assumption, highlighting that renters of independent houses face higher fire insurance premiums compared to renters of apartments in a condominium, even when the rent prices are the same.
We think this is because houses may not benefit from the shared fire protection measures present in a condominium, apartments in a condominium are usually built with fire safety features which can help mitigate the risk of fire and reduce insurance premiums.
Additionally, houses might require additional coverage for detached structures, such as garages, which can contribute to higher premiums compared to apartment units.

Moreover, we see some orange points situated in the blue line, these anomalies are likely errors in the data, and our analysis has enabled us to detect them.
We can use this to correct inaccuracies in the dataset allowing us to get more robust and accurate predictions.

## Conclusions

In conclusion, adopting a diversified approach by investing in both independent houses and apartments within condominiums can be a successful strategy to reach to a wider range of potential renters.
This allows the company to accommodate individuals with diverse preferences and priorities.
Some renters prioritize privacy, space, freedom, and are willing to choose independent houses despite their higher fire insurance premiums/rent prices and relatively distant locations from the city center.
On the other hand, there are those who prefer the affordability and security of apartment living, leading them to opt for apartment units within condominiums.
The company should evaluate potential returns on investments.
This assessment should be based on the location, investment costs, associated risks and the company's capital.
Our predictive model, which accurately estimates rent amounts, can serve as a valuable tool in this evaluation process.

# 7 - Clustering the houses for rental

Our second objective is to *Cluster the houses for rental according to their characteristics*; we are going to compare K-Means and Hierarchical Clustering choosing an optimal number k of clusters with the silhouette and elbow methods.
We're going to compare partitions and average silhouette width to determine which one best resembles reality, then evaluate agreement between the partitions and the different values in each cluster with respect to the initial dataset.

```{r fig71, fig.height = 3, fig.width = 7, fig.align = "center", warning=FALSE, message=FALSE, echo=FALSE}
# REMOVE CATEGORICAL FEATURES
Data_sc <- Data %>% select_if(is.numeric)
Data_sc <- as.data.frame(scale(Data_sc))

dist_p <- factoextra::get_dist(Data_sc,method = "pearson")
dist_e <- factoextra::get_dist(Data_sc,method = "euclidean")

# Optimal K
set.seed(444)

```

We get the following results for the best number of clusters silhouette wise:

```{r}
sil_k <- fviz_nbclust(
  Data_sc,
  FUNcluster = kmeans,
  diss = dist_e,
  method = "silhouette",
  print.summary = TRUE,
  k.max = 10
)
silh <-fviz_nbclust(
  Data_sc,
  FUNcluster = factoextra::hcut,
  diss = dist_e,
  method = "silhouette",
  print.summary = TRUE,
  k.max = 10
)
print(paste("Best k for K-Means(sil):",which(sil_k$data$y == max(sil_k$data$y)),", for HC(sil):",which(silh$data$y == max(silh$data$y))))

```

Consulting also the elbow method plot, we want to employ K-Means with k = 2 and Hierarchical with k = 4, both using euclidean distance as it proved to be more efficient for us.
We get the following partitions:

```{r fig73, fig.height = 3, fig.width = 5, fig.align = "center",echo=FALSE,warning=FALSE,message=FALSE}
# Trying with 2 clusters
# kmeans
km2 = kmeans(Data_sc, 2, nstart = 1, iter.max = 1e2)
kmv2 <- fviz_cluster(km2, data = Data_sc, geom = "point", 
             ggtheme = theme_minimal(), main = "K-Means")

# Evaluating silhouette
kmv2
sk2 = silhouette(km2$cluster, 
                 dist = dist_e)
sk2v <- fviz_silhouette(sk2,print.summary = FALSE)
print(paste("K-Means AVG Silhouette width:",mean(sk2v$data$sil_width),", totWSS:",km2$tot.withinss))
```

```{r fig74, fig.height = 3, fig.width = 5, fig.align = "center", echo=FALSE,warning=FALSE,message=FALSE}
# Hierarchical
hc2c <- factoextra::hcut(x = dist_e, 
                         k = 4,
                         hc_method = "ward.D2")
e2c <-factoextra::fviz_cluster(list(data = Data_sc, cluster = hc2c$cluster), main = "Hierarchical",labelsize = 0)

e2c # This plots partitions and silhouettes for both

print(paste("Hierarchical's AVG Silhouette width:",hc2c$silinfo$avg.width))
```

We're more convinced with the K-Means partition as it has higher average silhouette and partitions the datapoints clearly, while the hierarchical one has a lot of overlap between clusters, so we could reasonably say it has lower between-sum-of-squares and higher within-sum-of-squares.

By looking at the agreement index we can deduce more about the difference between partitions:

```{r,warning=FALSE,message=FALSE}
print(paste("Agreement betweenn K-Means and HC",mclust::adjustedRandIndex(km2$cluster,hc2c$cluster)))
```

The two partitions differ in the number of clusters and algorithms used, since we feel more confident about the 2-Means partition as it best represents our image of the data (i.e. lower end cheaper housing and higher end more expensive housing) we're going to rely on the K-Means partition for further analysis, so let's see what it represents with respect to the initial dataset:

```{r}
# Making 2 dfs for each cluster
clust1 <- Data[which(km2$cluster == 1),]
clust2 <- Data[which(km2$cluster == 2),]
print(paste("Average rent amount in: Cluster 1:",round(mean(clust1$rent.amount..R..),3),", Cluster 2: ",round(mean(clust2$rent.amount..R..),3)))
```

As we can see the average rent amount in cluster 2 is more than 3 times that of cluster 1, as we've mentioned before we think this reflects lower and higher end housing, but to come to more solid conclusions let's see some more data:

```{r}
print(paste("Percentage of houses in Cluster 1:",round(sum(as.numeric(summary(clust1$city))/nrow(Data)),3),", in Cluster 2:",round(sum(as.numeric(summary(clust2$city))/nrow(Data)),3)))

print(paste("Average property tax in Cluster 1:",round(mean(clust1$property.tax..R..),3),", in Cluster 2:",round(mean(clust2$property.tax..R..),3)))

print(paste("Average rent per room in Cluster 1:",round(mean(clust1$rent.amount..R../clust1$rooms),3),", in Cluster 2:",round(mean(clust2$rent.amount..R../clust2$rooms),3)))
```

## Pie chart for cities with respect to cluster

```{r fig000, fig.align='center',fig.width=8,fig.height=2}
display_city_pie_chart <- function(Data, data_name){  
  
  city_df <- as.data.frame(table(Data$city))
  colnames(city_df) <- c("city", "count")
  
  city_df$percentage <- city_df$count / sum(city_df$count)
  
  ggplot(city_df, aes(x = "", y = percentage, fill = city)) +
    geom_bar(width = 1, stat = "identity") +
    coord_polar("y", start = 0) + 
    scale_fill_brewer(palette = "Set3") +
    theme_void() +
    labs(title = paste("Pie Chart of Cities in", data_name), fill = "City")
}

pie1 <- display_city_pie_chart(clust1,"Cluster1")
pie2 <- display_city_pie_chart(clust2,"Cluster2")
ggarrange(pie1,pie2,nrow = 1,ncol = 2)

```

## Conclusions

A partition in higher and lower end housing shows dynamics of the region of brazil we have data about, where the majority of houses are located in Sao Paulo which is the most expensive place on average but also the most frequent city overall, which means that there can be both more and less affordable housing there.
The data also confirms the fact that tax and rent per room follow act similarly as rent in the different clusters.
As for our business' interests: If the company's objective is to acquire a larger volume of real estate with lower rent charges, they may consider focusing on the first cluster of houses.
These houses are likely to have lower acquisition costs, allowing the company to invest in a larger number of properties while keeping rental rates affordable.
This strategy would attract renters looking for more budget-friendly options.
On the other hand, if the company is more interested in high end properties that can command higher rental rates (despite higher acquisition costs), we would direct their attention towards the second cluster.
If the company aims to diversify their real estate portfolio, we can provide insights into potential acquisitions that vary in terms of cost.
We can identify properties that range from more affordable options to higher end investments.
This allows the company to consider a mix of properties, balancing their rental portfolio and appealing to a broader range of potential renters.
